---
title: "EMA single case data"
output: 
  github_document:
    toc: True
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
```{r echo = F, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)
library(lubridate)

```



```{r echo=FALSE}
## sanitise the EMA csv downloaded from M-path before importing it here.
# use th bash script "sanitiseEMA.sh" in ../data/
# Read the CSV data
data <- read.csv("C:/Documents and Settings/d.vitali/Desktop/Github/CRIISP-WP6/Data/EMA_basic_sanitized.csv", na.strings = "",sep = ";")

# select participant alias
data_1c<-data %>% filter(alias=="P005")

# Remove columns that are all NA
data_1c <- data_1c %>%
  select(where(~!all(is.na(.))))

```


# Case details

  * Sex: F
  * Age: 32

## BASELINE DATA:

### BPI

 The BPI allows patients to rate the severity of their pain and the degree to which their pain interferes with common dimensions of feeling and function. The interference scale includes  seven domains: 1. General activity,  2. Mood,  3. Walking ability,  4. Normal work (including housework),  5. Relations with other people,  6. Sleep, and  7. Enjoyment of life.

**Pain Location**

  * Lower Back 
  * Right hand
  * Left Glute
  * Left leg

**Pain intensity in the previous 24h:**

|        | score   |
|--------|---------|
| worst  | 10/10   |
| average| 6.75/10 |

**Pain interference in the previous 24h:**

|        | score   |
|--------|---------|
|average | 7.85/10 |
 
  
**Pain Medication**

  * Amitriptyline (tricyclic antidepressant)
  * duloxetine  (SNRI)

  * codeine (opioid medication)
  * naproxen (NSAID)


### Psychological wellbeing

#### PHQ9
The PHQ-9 is a 9 item questionnaire aimed at scoring each of the nine DSM-IV criteria. It is generally used to monitor the severity of depression symptomatology and considers scores above 20 as *severe* (range 0:27).

|        | score |
|--------|-------|
|**PHQ9**| 24/27 |

#### PSEQ
The Pain Self-Efficacy Questionnaire (PSEQ) is a 10-item questionnaire developed to assess the confidence people with ongoing pain have in performing activities while in pain. The scoring range is 0:60 with higher scores indicating greater confidence in dealing with pain. Scores of 40 (percentile 50) are generally associated with favourable chance of "returning to work", lower scores (for example a raw score of 30, percentile = 18) tend to predict less sustainable gains.

The PSEQ scores are helpful in developing a formulation around psychological factors (e.g. beliefs) that influence the respondent's reaction to injury and pain.

|        | score |
|--------|-------|
|**PSEQ**| 16/60 |



# data prep

```{r echo = F, warnings= F}
# Convert Unix timestamp to datetime
data_1c<-data_1c %>%
  mutate(
    datetime = as_datetime(timeStampStart),
    date = date(datetime),
    Mood_dec = Mood_smiley/10
  )

data_1c <- data_1c %>%
  arrange(datetime) %>%
  mutate(day_count = as.numeric(date - min(date, na.rm = TRUE)),
    is_weekend = wday(date) %in% c(1, 7)
    )
```

# Reported pain locations
```{r echo = F, warnings= F}
##################################
#################################
# Split the pain locations into separate rows
pain_locations <- data_1c %>%
  select(datetime, Pain.location_bodyParts) %>%
  separate_rows(Pain.location_bodyParts, sep = ",") %>%
  mutate(Pain.location_bodyParts = trimws(Pain.location_bodyParts))

# Overall frequency of pain locations
pain_freq <- pain_locations %>%
  count(Pain.location_bodyParts, sort = TRUE) %>%
  mutate(percentage = n/sum(n)*100)

# Print overall frequencies
#print("Overall Pain Location Frequencies:")
#print(pain_freq)

# Create a bar plot of pain locations
ggplot(pain_freq, aes(x = reorder(Pain.location_bodyParts, n), y = n)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Frequency of Pain Locations",
       x = "Location",
       y = "Count") +
  theme_minimal()
```

## Daily Pain Locations and flares-up
```{r echo = F, warnings= F}

location_order <- c(
  "left foot", "right foot",
  "left knee", "right knee",
  "left leg", "right leg",
  "back", "belly",
  "left hand", "left arm",
  "right hand","right arm",
  "neck and shoulders", "neck", "head"
)

# Split pain locations and create weighted intensity
daily_pain_weighted <- data_1c %>%
  select(day_count, Pain.location_bodyParts, Pain.intensity_sliderNeutralPos) %>%
  separate_rows(Pain.location_bodyParts, sep = ",") %>%
  mutate(
    Pain.location_bodyParts = trimws(Pain.location_bodyParts),
    # Convert to factor with custom order
    Pain.location_bodyParts = factor(Pain.location_bodyParts, 
                                   levels = location_order,
                                   ordered = TRUE)
  ) %>%
  group_by(day_count, Pain.location_bodyParts) %>%
  summarise(
    frequency = n(),
    avg_intensity = mean(Pain.intensity_sliderNeutralPos, na.rm = TRUE),
    weighted_score = frequency * avg_intensity
  ) %>%
  ungroup()

# Create weighted heatmap
p1<-ggplot(daily_pain_weighted, aes(x = day_count, y = Pain.location_bodyParts, fill = weighted_score)) +
  geom_tile() +
  scale_fill_gradient(low = "lightgrey", high = "red") +
  scale_x_continuous(breaks = function(x) seq(floor(min(x)), ceiling(max(x)), by = 1)) +
  labs(
    title = "Daily Pain Locations and flares-up",
    x = "Days from Start",
    y = "Location",
    fill = "Pain Score\n(Frequency × Intensity)"
  ) +
  theme_minimal()
p1
```

# Daily Mood, Stress, and pain intensity

```{r echo = F, warning = F}
p2<-ggplot(data_1c, aes(x = day_count)) +
  # Add horizontal line for mood inflection point
  geom_hline(yintercept = 5, linetype = "dashed", color = "gray50", alpha = 0.5) +
  # Smooth lines with less smoothening
  geom_smooth(aes(y = Mood_dec, color = "Mood"), 
              method = "loess", span = 0.2, se = FALSE) +
  geom_smooth(aes(y = Stress_sliderNeutralPos, color = "Stress"), 
              method = "loess", span = 0.2, se = FALSE) +
  geom_smooth(aes(y = Pain.intensity_sliderNeutralPos, color = "Pain"), 
              method = "loess", span = 0.2, se = FALSE) +
  # Add points for actual data
  geom_point(aes(y = Mood_dec, color = "Mood"), alpha = 0.3) +
  geom_point(aes(y = Stress_sliderNeutralPos, color = "Stress"), alpha = 0.3) +
  geom_point(aes(y = Pain.intensity_sliderNeutralPos, color = "Pain"), alpha = 0.3) +
  scale_color_manual(values = c("Mood" = "blue", "Stress" = "red", "Pain" = "green")) +
  scale_y_continuous(breaks = 0:10, limits = c(0,10)) +
  scale_x_continuous(breaks = function(x) seq(floor(min(x)), ceiling(max(x)), by = 1)) +
  labs(title = "Mood, Stress, and Pain intensity Over Time",
       y = "Score",
       x = "Days from Start",
       caption = "Low mood < 5, Good mood > 5") +
  theme_minimal()

p2
```


## Daily pain interference
```{r echo = F, warning= F}
p3<-ggplot(data_1c, aes(x = day_count)) +
  # Add horizontal line for mood inflection point
  geom_hline(yintercept = 5, linetype = "dashed", color = "gray50", alpha = 0.5) +
  # Smooth lines with less smoothening
  geom_smooth(aes(y = Pain.interf..plans_sliderNeutralPos, color = "Interference with plans"), 
              method = "loess", span = 0.4, se = FALSE) +
  geom_smooth(aes(y = Pain.interf..Life.enjoyment_sliderNeutralPos, color = "Interference with life enjoiment"), 
              method = "loess", span = 0.4, se = FALSE) +
  geom_smooth(aes(y = Pain.intensity_sliderNeutralPos, color = "Pain intensity"), 
              method = "loess", span = 0.3, se = FALSE) +
   geom_smooth(aes(y = Stress_sliderNeutralPos, color = "Stress"), 
              method = "loess", span = 0.4, se = FALSE) +
  # Add points for actual data
  geom_point(aes(y = Pain.interf..plans_sliderNeutralPos, color = "Interference with plans"), alpha = 0.3) +
  geom_point(aes(y = Pain.interf..Life.enjoyment_sliderNeutralPos, color = "Interference with life enjoiment"), alpha = 0.3) +
  geom_point(aes(y = Pain.intensity_sliderNeutralPos, color = "Pain intensity"), alpha = 0.3) +
  geom_point(aes(y = Stress_sliderNeutralPos, color = "Stress"), alpha = 0.3) +

   scale_color_manual(values = c("Interference with plans" = "green", 
                                "Interference with life enjoiment" = "black", 
                                "Pain intensity" = "red",
                                "Stress" = "blue")) +
  scale_y_continuous(breaks = 0:10, limits = c(0,10)) +
  scale_x_continuous(breaks = function(x) seq(floor(min(x)), ceiling(max(x)), by = 1)) +
  labs(title = "",
       y = "Score",
       x = "Days from Start") +
  theme_minimal()

p3
```

## Pain Mood and Pain interference Over Time

```{r echo = F, warning= F, fig.width=10, fig.height=10}
library(patchwork)
# Combine plots vertically
combined_plot <- p1 / p2 / p3 +
  plot_layout(heights = c(1, 1, 1)) +
  plot_annotation(
    title = "Pain Mood and Pain interference Over Time",
    theme = theme_minimal()
  )

# Display the combined plot
combined_plot

```

## Social Activity
```{r echo = F, warning= F, fig.width=10, fig.height=7}

# Create a function to process the data with a type indicator
create_daily_data <- function(data, column_name, type) {
  data %>%
    select(day_count, !!sym(column_name)) %>%
    separate_rows(!!sym(column_name), sep = ",") %>%
    mutate(
      activity = trimws(!!sym(column_name)),
      type = type
    ) %>%
    group_by(day_count, activity, type) %>%
    summarise(count = n(), .groups = 'drop') %>%
    filter(!is.na(activity) & activity != "")
}

# Create datasets for each type
social_data <- create_daily_data(data_1c, "Social_multipleChoice_string", "Social Contacts")
location_data <- create_daily_data(data_1c, "Location_multipleChoice_string", "Locations")
activity_data <- create_daily_data(data_1c, "Activity_multipleChoice_string", "Activity")

# Combine all data
combined_data <- bind_rows(activity_data, social_data, location_data)

combined_data <- combined_data %>%
  mutate(
    activity = case_when(
      activity == "Active free-time (walking/ sports/ gardening)" ~ "Active free-time",
      activity == "Non-active free-time (TV/computer/ video games/ social media/ reading)" ~ "Non-active free-time",
      activity == "Socialising outside (going to the pub/ cinema/ theatre/ going out)" ~ "Socialising outside",
      TRUE ~ activity
    )
  )

# Create the combined plot
p4<-ggplot(combined_data, aes(x = day_count, y = activity, fill = count)) +
  geom_tile() +
  scale_fill_gradient(low = "lightgrey", high = "red") +
  scale_x_continuous(breaks = function(x) seq(floor(min(x)), ceiling(max(x)), by = 1)) +
  facet_grid(type ~ ., scales = "free_y", space = "free_y") +
  labs(
    title = "Daily Activities, Social Contacts, and Locations",
    x = "Days from Start",
    y = "",
    fill = "Frequency"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    strip.text.y = element_text(angle = 0),
    panel.spacing = unit(1, "lines"),
    legend.position = "right"
  )

p5<-ggplot(data_1c, aes(x = day_count)) +
  # Add horizontal line for mood inflection point
  geom_hline(yintercept = 5, linetype = "dashed", color = "gray50", alpha = 0.5) +
  # Smooth lines with less smoothening
  geom_smooth(aes(y = Social.support_sliderNeutralPos, color = "Social Support"), 
              method = "loess", span = 0.3, se = FALSE) +
   geom_smooth(aes(y = Stress_sliderNeutralPos, color = "Stress"), 
              method = "loess", span = 0.3, se = FALSE) +
  geom_smooth(aes(y = Mood_dec, color = "Mood"), 
              method = "loess", span = 0.3, se = FALSE) +
  # Add points for actual data
  geom_point(aes(y = Social.support_sliderNeutralPos, color = "Social Support"), alpha = 0.3) +
  geom_point(aes(y = Stress_sliderNeutralPos, color = "Stress"), alpha = 0.3) +
  geom_point(aes(y = Mood_dec, color = "Mood"), alpha = 0.3) +

  scale_color_manual(values = c("Social Support" = "black",
                                "Mood" = "blue",
                                "Stress" = "red")) +
  scale_y_continuous(breaks = 0:10, limits = c(0,10)) +
  scale_x_continuous(breaks = function(x) seq(floor(min(x)), ceiling(max(x)), by = 1)) +
  labs(title = "",
       y = "Score",
       x = "Days from Start") +
  theme_minimal()



# Combine plots vertically with adjusted heights
combined_plot <- p4 / p5 + 
  plot_layout(heights = c(4, 1)) +  # Adjust the ratio between plots (2:1)
  plot_annotation(
    title = "Social Interactions Analysis",
    theme = theme_minimal()
  )

combined_plot

```
