---
title: "EMA single case data"
output: 
  github_document:
    toc: True
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
```{r echo = F}
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(zoo)
library(patchwork)


source("C:/Users/d.vitali/Desktop/Github/CRIISP-WP6/Sensors-API/Withings-API/withings_func.R")

```

set participant data folder
```{r warning = F, fig.width=10, fig.height=5}
#folder_name<-"data_ART_1727097278"
#  start_date<-"2024-09-20 06:00:00"
#  end_date  <-"2024-09-25 06:00:00"

folder_name<-"data_PIP_1732759873"
  start_date<-"2024-10-22 06:00:00"
  end_date  <-"2024-11-23 06:00:00"

#folder_name<-"data_JAA_1727098304"
#  start_date<-"2024-09-04 06:00:00"
#  end_date  <-"2024-09-12 06:00:00"

#folder_name<-"data_CEC_1727186111"
#  start_date<-"2024-08-29 06:00:00"
#  end_date  <-"2024-09-08 06:00:00"




path<-paste0("C:/Users/d.vitali/Desktop/Github/CRIISP-WP6/Data/",folder_name,"/")


```

# BODY temperature
monitored  ~ 1400/day (~58/hr)

```{r echo = F,  warning = F, fig.width=10, fig.height=5}
# Sample data
temperature_csv <- read.csv(file = paste0(path,"raw_core_body_temperature_Core body temperature.csv"), header = TRUE, sep = ",")

# Convert start column to datetime
temperature_csv$start <- ymd_hms(temperature_csv$start,  tz = "UTC")

temperature_data<-expand_csv_data(temperature_csv) %>%
  filter(timestamp < end_date & timestamp > start_date)

temperature_data %>%
  # Create a new column 'day' that groups by date
  mutate(day = as.Date(timestamp)) %>%
  # Group by day
  group_by(day) %>%
  # Count non-NA values in the heart_rate column
  summarise(non_na_count = sum(!is.na(value))) 
```

```{r echo = F,  warning = F, fig.width=10, fig.height=5}
# First, let's create a function for consistent plotting
plot_24h_temp <- function(data, start_date) {
  end_date <- start_date + days(1)

  data %>% 
    filter(timestamp >= start_date, timestamp < end_date) %>%
    ggplot() +
    geom_line(aes(x = timestamp, y = value), color = "blue") +
    scale_y_continuous(
      name = "Temperature") +
    scale_x_datetime(date_labels = "%H:%M", date_breaks = "1 hours") +
    labs(
      title = paste("Temperature Changes:", format(start_date, "%Y-%m-%d")),
      x = "Time"
    ) +
    theme_minimal()
}

# Get the start and end dates from your data
start_date <- floor_date(min(temperature_data$timestamp), "day")
end_date <- ceiling_date(max(temperature_data$timestamp), "day")
dates_seq <- seq(start_date, end_date - days(1), by = "1 day")

# Create a list of plots
plots <- lapply(dates_seq, function(d) plot_24h_temp(temperature_data, d))

# Print each plot
#for(p in plots) {
#  print(p)
#}


```
##Temperature cleaning

Temperature changes are key indicators of sleep/wake transitions 
 - Temperature naturally drops during sleep onset 2,5
 - Warmer temperatures accompany waking up from being asleep 3

data cleaning should suit this:

1. Detect temperature change patterns 5
3. detect and correct drifts and spikes Using a single rolling mean but accounts for natural temperature variations throughout sleep/awake  cycle 


```{r echo = F,  warning = F, fig.width=10, fig.height=5}
cleaned_temperature_data <- temperature_data %>%
  arrange(timestamp) %>%
  mutate(
    hour = lubridate::hour(timestamp),
    
    # Basic time classification
    likely_awake = hour >= 8 & hour < 23,
    
    # First pass: Remove physiologically impossible values
    value_filtered = if_else(value > 38.5, NA_real_, value),
    
    # Calculate temperature changes
    temp_change = c(0, diff(value_filtered)),
    rolling_temp_change = rollapply(temp_change, width = 31, 
                                    FUN = mean, na.rm = TRUE, fill = "extend"),
    
    # Calculate rolling mean for reference
    rolling_mean = rollapply(value_filtered, width = 361, 
                             FUN = mean, na.rm = TRUE, fill = "extend"),
    
    # Calculate local variability
    local_mad = rollapply(value_filtered, width = 181, 
                          FUN = function(x) median(abs(x - median(x, na.rm = TRUE)), na.rm = TRUE),
                          fill = "extend"),
    
    # Classify sleep/wake based on time and temperature patterns
    is_sleep = case_when(
      # During night hours (23:00-08:00)
      !likely_awake & rolling_temp_change < 0 ~ TRUE,  # Temperature dropping
      !likely_awake & value_filtered < rolling_mean ~ TRUE,  # Lower than average
      # During day hours
      likely_awake & rolling_temp_change > 0 ~ FALSE,  # Temperature rising
      likely_awake ~ FALSE,  # Default to awake during day hours
      TRUE ~ !likely_awake  # Default to time-based classification
    ),
    
    # Apply corrections
    cleaned_value = case_when(
      # Remove physiologically impossible values
      is.na(value_filtered) ~ rolling_mean,
      
      # Correct spikes
      abs(value_filtered - rolling_mean) > (2 * local_mad) ~ rolling_mean,
      
      # Correct drift
      abs(value_filtered - rolling_mean) > (1.5 * local_mad) ~
        (0.7 * rolling_mean + 0.3 * value_filtered),
      
      # Keep original valid values
      TRUE ~ value_filtered
    ),
    
    # Final smoothing
    final_value = rollapply(cleaned_value, width = 5, 
                            FUN = median, na.rm = TRUE, fill = "extend")
  )



ggplot(cleaned_temperature_data) +
  geom_rect(aes(xmin = timestamp, xmax = lead(timestamp), 
                ymin = -Inf, ymax = Inf, fill = is_sleep), 
            alpha = 0.2) +
  geom_line(aes(x = timestamp, y = value), alpha = 0.7, color = "orange") +
  geom_line(aes(x = timestamp, y = final_value), color = "red", size = 0.5) +
#  geom_line(aes(x = timestamp, y = rolling_temp_change * 10 + mean(value, na.rm = TRUE)), 
#            color = "blue", alpha = 0.3) +  # Scaled temperature change
  scale_fill_manual(values = c("white", "gray"), 
                    labels = c("Awake", "Sleep")) +
  labs(title = "Temperature with Dynamic Corrections",
       subtitle = "Gray: Original | Red: Corrected | Blue: Temperature Change (scaled)",
       x = "Time", y = "Temperature") +
  theme_minimal()


# Create histogram with density plots
ggplot(cleaned_temperature_data, aes(x = value_filtered)) +
  geom_histogram(aes(fill = likely_awake), 
                 position = "identity",
                 alpha = 0.5,
                 bins = 50) +
  geom_density(aes(y = ..count.. * 2, color = likely_awake),  # Scale density to match histogram
               alpha = 0.2) +
  scale_fill_manual(values = c("darkblue", "orange"),
                   labels = c("Night (23:00-08:00)", "Day (08:00-23:00)")) +
  scale_color_manual(values = c("darkblue", "orange"),
                    labels = c("Night (23:00-08:00)", "Day (08:00-23:00)")) +
  labs(title = "Distribution of Body Temperature by Time of Day",
       subtitle = "Showing bimodal pattern between day and night temperatures",
       x = "Temperature",
       y = "Count",
       fill = "Time Period",
       color = "Time Period") +
  theme_minimal() +
  theme(legend.position = "top")

# Add summary statistics
summary_stats <- cleaned_temperature_data %>%
  group_by(likely_awake) %>%
  summarise(
    mean_temp = mean(value_filtered, na.rm = TRUE),
    sd_temp = sd(value_filtered, na.rm = TRUE),
    median_temp = median(value_filtered, na.rm = TRUE),
    q25 = quantile(value_filtered, 0.25, na.rm = TRUE),
    q75 = quantile(value_filtered, 0.75, na.rm = TRUE)
  )

print(summary_stats)

```

```{r echo = F,  warning = F, fig.width=10, fig.height=5}
temperature_10min <- cleaned_temperature_data %>%
  # Round timestamp to nearest 10 minutes
  mutate(
    timestamp_10min = floor_date(timestamp, unit = "10 minutes")
  ) %>%
  # Group by 10-minute intervals and calculate mean
  group_by(timestamp_10min) %>%
  summarise(
    value = mean(final_value, na.rm = TRUE)
  ) %>%
  # Remove any NaN that might result from all NA values in a window
  filter(!is.nan(value))
```
# STEPS
ad hoc (whenever more than 0 step is recorded)
```{r echo = F,  warning = F, fig.width=10, fig.height=5}
# Sample data
steps_csv <- read.csv(file = paste0(path,"raw_tracker_steps.csv"), header = TRUE, sep = ",")

# Convert start column to datetime
steps_csv$start <- ymd_hms(steps_csv$start,  tz = "UTC")

steps_data<-expand_csv_data(steps_csv) %>%
  filter(timestamp < end_date & timestamp > start_date)

steps_data %>%
  # Create a new column 'day' that groups by date
  mutate(day = as.Date(timestamp)) %>%
  # Group by day
  group_by(day) %>%
  # Count non-NA values in the heart_rate column
  summarise(non_na_count = sum(!is.na(value)))

```



# Heart Rate
monitored  ~ 140/day (~6/hr)

```{r echo = F,  warning = F, fig.width=10, fig.height=5}
# Sample data
HR_csv <- read.csv(file = paste0(path,"raw_hr_hr.csv"), header = TRUE, sep = ",")

# Convert start column to datetime
HR_csv$start <- ymd_hms(HR_csv$start,  tz = "UTC")

HR_data<-expand_csv_data(HR_csv) %>%
  filter(timestamp < end_date & timestamp > start_date)


HR_data %>%
  # Create a new column 'day' that groups by date
  mutate(day = as.Date(timestamp)) %>%
  # Group by day
  group_by(day) %>%
  # Count non-NA values in the heart_rate column
  summarise(non_na_count = sum(!is.na(value)))

```

# Activity Intensity
monitored  ~ 200/day (~8/hr)
```{r echo = F,  warning = F, fig.width=10, fig.height=5}
# Sample data
activity_intensity_csv <- read.csv(file = paste0(path,"raw_tracker_Activity Intensity.csv"), header = TRUE, sep = ",")

# Convert start column to datetime
activity_intensity_csv$start <- ymd_hms(activity_intensity_csv$start,  tz = "UTC")

activity_intensity_data<-expand_csv_data(activity_intensity_csv) %>%
  filter(timestamp < end_date & timestamp > start_date)

activity_intensity_data$value[activity_intensity_data$value>5]<-NA

activity_intensity_data%>%
  # Create a new column 'day' that groups by date
  mutate(day = as.Date(timestamp)) %>%
  # Group by day
  group_by(day) %>%
  # Count non-NA values in the heart_rate column
  summarise(non_na_count = sum(!is.na(value)))

```

```{r echo = F,  warning = F, fig.width=10, fig.height=5}
# Resample activity intensity data to 10-minute averages
activity_intensity_10min <- activity_intensity_data %>%
  # Round timestamp to nearest 10 minutes
  mutate(
    timestamp_10min = floor_date(timestamp, unit = "10 minutes")
  ) %>%
  # Group by 10-minute intervals and calculate mean
  group_by(timestamp_10min) %>%
  summarise(
    value = mean(value, na.rm = TRUE)
  ) %>%
  # Remove any NaN that might result from all NA values in a window
  filter(!is.nan(value))

```

# Sleep states
```{r echo = F,  warning = F, fig.width=10, fig.height=5}
# Sample data
sleepstate_csv <- read.csv(file = paste0(path,"raw_tracker_sleep-state.csv"), header = TRUE, sep = ",")

# Convert start column to datetime
sleepstate_csv$start <- ymd_hms(sleepstate_csv$start,  tz = "UTC")

sleepstate_data<-expand_csv_data(sleepstate_csv) %>%
  filter(timestamp < end_date & timestamp > start_date)

sleepstate_data %>%
  # Create a new column 'day' that groups by date
  mutate(day = as.Date(timestamp)) %>%
  # Group by day
  group_by(day) %>%
  # Count non-NA values in the heart_rate column
  summarise(non_na_count = sum(!is.na(value)))

```

## Sleep summary
```{r echo = F,  warning = F, fig.width=10, fig.height=5}
# Sample data
sleepsummary_data <- read.csv(file = paste0(path,"sleep.csv"), header = TRUE, sep = ",")

# Convert start column to datetime
sleepsummary_data$from <- ymd_hms(sleepsummary_data$from,  tz = "UTC")

sleepsummary_data$to <- ymd_hms(sleepsummary_data$to,  tz = "UTC")

sleepsummary_data<-sleepsummary_data %>%
  filter(from < end_date & from > start_date)


```

# ECG
```{r echo = F,  warning = F, fig.width=10, fig.height=5}
# Sample data
ECG_csv <- read.csv(file = paste0(path,"signal.csv"), header = TRUE, sep = ",")

# Convert date column to datetime
ECG_csv$date <- ymd_hms(ECG_csv$date,  tz = "UTC") 

ECG_data<-as_tibble(ECG_csv)%>%
  filter(date < end_date & date > start_date)


```


# Merge steps, activity, hr, temperature
```{r echo = F,  warning = F, fig.width=10, fig.height=5}

act_data <- full_join(temperature_10min, activity_intensity_10min, by = "timestamp_10min")

colnames(act_data)<-c("timestamp","temp","activity_int")

act_data <- full_join(act_data, HR_data, by = "timestamp")

colnames(act_data)<-c("timestamp","temperature","activity_int","hr")


```


# Sleep detection

```{r echo = F,  warning = F, fig.width=10, fig.height=5}

plot_data<-act_data

# Reshape the data for faceting
act_data_long <- plot_data %>%
  pivot_longer(cols = c(activity_int, temperature, hr),
               names_to = "metric",
               values_to = "value")

# Prepare sleep periods from summary data
sleep_bands <- sleepsummary_data %>%
  select(from, to) %>%
  mutate(source = "sleep_summary")

# Create the main plot
ggplot(act_data_long %>%
         mutate(
           value = case_when(
             metric == "temperature" & (value < 35 | value > 45) ~ NA_real_,
             TRUE ~ value
           ),
           # Create day numbers starting from 0
           day_number = as.numeric(difftime(timestamp, min(timestamp, na.rm = TRUE), units = "days"))
         ), 
       aes(x = day_number, y = value)) +  # Change x aesthetic to day_number
  geom_rect(data = sleep_bands %>%
              mutate(day_from = as.numeric(difftime(from, min(act_data_long$timestamp, na.rm = TRUE), units = "days")),
                     day_to = as.numeric(difftime(to, min(act_data_long$timestamp, na.rm = TRUE), units = "days"))),
            aes(xmin = day_from, 
                xmax = day_to),
            ymin = -Inf, 
            ymax = Inf,
            fill = "gray80",
            alpha = 0.4,
            inherit.aes = FALSE) +
  # Separate geom_point layers for different metrics with different alphas
  geom_point(aes(color = metric), alpha = 0.2) +
  scale_color_manual(values = c(
    "activity_int" = "blue",
    "steps" = "black",
    "temperature" = "red",
    "hr" = "red"
  )) +

  # Add breaks for each day
  scale_x_continuous(breaks = function(x) seq(0, floor(max(x)), by = 1)) +

  facet_wrap(~ metric, scales = "free_y", nrow = 4) +

  labs(title = paste(folder_name, "Activity Intensity, Heart Rate, Steps and Temperature Over Time"),
       subtitle = "Gray bands indicate sleep periods",
       x = "Days from Start",
       y = "Value",
       color = "Metrics") +

  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.spacing = unit(1, "lines"),
    legend.position = "right"
  )



```

#MISSINGNESS

## explore missing hr datapoints
```{r echo = F,  warning = F, fig.width=10, fig.height=5}
act_data_NA<-act_data %>% filter(!is.na(hr))
act_data_NA <- act_data_NA %>%
  arrange(timestamp) %>%  
  mutate(minutes_diff = as.numeric(difftime(timestamp, lag(timestamp), units = "mins")))

ggplot(act_data_NA, aes(x = timestamp, y = minutes_diff)) +
  geom_line() +
  scale_y_continuous(breaks = seq(0, max(act_data_NA$minutes_diff, na.rm = TRUE), by = 10)) + 
  labs(title = "Time Difference Between HR Observations", 
       x = "Timestamp", 
       y = "Minutes Difference") +
  theme_minimal()


```

## explore missing Temperature datapoints
```{r echo = F,  warning = F, fig.width=10, fig.height=5}
act_data_NA<-act_data %>% filter(!is.na(temperature))
act_data_NA <- act_data_NA %>%
  arrange(timestamp) %>%  
  mutate(minutes_diff = as.numeric(difftime(timestamp, lag(timestamp), units = "mins")))

ggplot(act_data_NA, aes(x = timestamp, y = minutes_diff)) +
  geom_line() +
  scale_y_continuous(breaks = seq(0, max(act_data_NA$minutes_diff, na.rm = TRUE), by = 10)) + 
  labs(title = "Time Difference Between Temp Observations", 
       x = "Timestamp", 
       y = "Minutes Difference") +
  theme_minimal()


```





# OTHER derived
```{r echo = F,  warning = F, fig.width=10, fig.height=5}
# Sample data
distance_csv <- read.csv(file = paste0(path,"raw_tracker_distance.csv"), header = TRUE, sep = ",")

# Convert start column to datetime
distance_csv$start <- ymd_hms(distance_csv$start,  tz = "UTC")

distance_data<-expand_csv_data(distance_csv)

```


```{r echo = F,  warning = F, fig.width=10, fig.height=5}
# Sample data
elevation_csv <- read.csv(file = paste0(path,"raw_elevation_Elevation.csv"), header = TRUE, sep = ",")

# Convert start column to datetime
elevation_csv$start <- ymd_hms(elevation_csv$start,  tz = "UTC")


elevation_data<-expand_csv_data(elevation_csv)
```


raw_tracker_ACTIREC_FEAT.csv
```{r echo = F,  warning = F, fig.width=10, fig.height=5}
# Sample data
actirec_csv <- read.csv(file = paste0(path,"raw_tracker_ACTIREC_FEAT.csv"), header = TRUE, sep = ",")

# Convert start column to datetime
actirec_csv$start <- ymd_hms(actirec_csv$start,  tz = "UTC")


actirec_data<-expand_csv_data(actirec_csv)


act_data <- full_join(act_data, actirec_data, by = "timestamp")

colnames(act_data)<-c("timestamp","temperature","activity_int","hr","steps")


plot_data<-act_data %>% filter(timestamp> start_date)

# Reshape the data for faceting
act_data_long <- plot_data %>%
  pivot_longer(cols = c(activity_int, steps, temperature, hr),
               names_to = "metric",
               values_to = "value")

# Create the plot with facets
ggplot(act_data_long, aes(x = timestamp, y = value)) +
  
  geom_point(data = act_data_long %>% filter(metric == "activity_int"), 
           aes(fill = metric)) +
  
  geom_point(data = act_data_long %>% filter(metric == "steps"), 
           aes(fill = metric)) +
  
  geom_point(data = act_data_long %>% filter(metric == "temperature"), 
            aes(color = metric)) +
  
  geom_point(data = act_data_long %>% filter(metric == "hr"), 
             aes(color = metric)) +
  
  # Labels and theme
  labs(title = paste(folder_name,"activity intensity, Heart Rate, steps, temperature Over Time"),
       x = "Time",
       y = "Value",
       color = "Metrics",
       fill = "Metrics") +
  
  # Facet by metric
  facet_wrap(~ metric, scales = "free_y", nrow = 4) +
  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```