---
title: "withingsAPI"
output: 
  github_document:
    toc: True
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r echo=F}
library("tidyr")
library("dplyr")
library("httr")
library("jsonlite")
```

```{r}
APIendpoint<-"https://wbsapi.withings.net"
AccessURL<-"https://account.withings.com/oauth2/token"
AuthURL <- "https://account.withings.com/oauth2_user/authorize2"
TokenURL<- "https://wbsapi.withings.net/v2/oauth2"
CallBackURL<-"http://localhost:1410/"
```

set working directory
```{r, echo=FALSE}
setwd("C:/Users/d.vitali/Desktop/Github/CRIISP-WP6/Sensors-API/Withings-API/")
```


# Load Necessary functions
```{r}
# this should read withings_func.R which contains a set of functions required in the pipeline
source("withings_func.R")

```

# Define Participant ID

```{r}
#set credentials for this participant
Participant_id<-"credentials/P_03.conf"
  
```


# Load Credentials
```{r}

source(Participant_id)
# this should read client_id, client_secret, and refresh_token as environment variables
```


# API setup

The browser will now open the participant details recorded on Withings 
(you may need to log in).

Please check that you are logged in as the participant for which you 
want to get API authorization codes.
 

```{r}
#   browseURL("https://www.withings.com/uk/en/account/profil")
```

```{r}
# Main logic to check if refresh_token is empty and handle accordingly
if (refresh_token == "") {
  cat("refresh_token is empty. Request a new token \n")
} else {
  cat("refresh_token found. Refreshing existing token.\n")
  tokens <- refresh_existing_token(refresh_token)
}
```

```{r}

# tokens <- request_new_tokens()
```


```{r}
# Access tokens can now be used for further API calls
access_token <- tokens$access_token
refresh_token <- tokens$refresh_token

```


```{r}

req <- httr::GET("https://wbsapi.withings.net/v2/user?action=getdevice", 
                 query = list(access_token = access_token))

httr::stop_for_status(req)
out <- httr::content(req, as = "text", encoding = "utf-8")
out <- jsonlite::fromJSON(out)

out$status

devices.df<-out$body$devices

devices.df$first_session_date <- as.POSIXct(devices.df$first_session_date, origin="1970-01-01") %>% format("%Y-%m-%d %H:%M:%S")
devices.df$last_session_date <- as.POSIXct(devices.df$last_session_date, origin="1970-01-01") %>% format("%Y-%m-%d %H:%M:%S")

deviceid<-out$body$devices$deviceid[1] # just in case I need to filter by deviceid


```
# SET start date enddate

Set this to three arbitrary days where I know there is data
```{r}

startdate   <-as.POSIXct(devices.df$first_session_date, origin="1970-01-01") %>% format("%Y-%m-%d")
enddate     <-as.POSIXct(devices.df$last_session_date, origin="1970-01-01") %>% format("%Y-%m-%d")
```

transform the time format of "startdate" in the format required by the different API calls
```{r}
## set time variable as unixtime
UnixStartDate <- as.POSIXct(paste(startdate, "00:00:00"), format = "%Y-%m-%d %H:%M:%S", tz = "UTC")
UnixStartDate<- as.numeric(unclass(UnixStartDate))

UnixEndDate <- as.POSIXct(paste(enddate, "23:59:59"), format = "%Y-%m-%d %H:%M:%S", tz = "UTC")
UnixEndDate <- as.numeric(unclass(UnixEndDate))

## set time variable as timezone
startdate_tz<-convert_date(startdate, "Date", tz)
enddate_tz<-convert_date(enddate, "Date", tz)

```

#API CALLS

## getactivity()

```{r}
##
data_fields <-"steps,distance,elevation,soft,moderate,intense,active,calories,hr_average,hr_min,hr_max,hr_zone_0,hr_zone_1,hr_zone_2,hr_zone_3"

req <- httr::GET("https://wbsapi.withings.net/v2/measure?action=getactivity", 
                 query = list(access_token = access_token,
                              startdateymd = startdate_tz, 
                              enddateymd = enddate_tz,
                              data_fields = data_fields,
                              offset = ""))
httr::stop_for_status(req)
out <- httr::content(req, as = "text", encoding = "utf-8")
out <- jsonlite::fromJSON(out)

out$status


activity.df<-out$body$activities

#write.csv(activity.df,"c:/Documents and Settings/d.vitali/Downloads/getactivity.csv")


```






## getintradayactivity()


```{r}

data_fields <-"steps,elevation,calories,distance,duration,heart_rate,spo2_auto"

# Fetch the intraday activity data for the time period
intra_day_activity_data <- fetch_ALL_intraday_activity(startdate, enddate, access_token, data_fields)


```

### HR and general movement data
```{r}

library(lubridate)

binned_data <- intra_day_activity_data %>%
  mutate(bin_min = floor_date(timestamp, "5 minutes")) %>%
  group_by(bin_min) %>%
  summarise(heart_rate = mean(heart_rate, na.rm = TRUE),   # Average heart rate in each 10-minute bin
            steps = sum(steps, na.rm = TRUE))              # Sum steps in each 10-minute bin


ggplot(binned_data, aes(x = bin_min)) +
  
  # point plot for heart rate
  geom_point(aes(y = heart_rate, color = "Heart Rate"), size = 1) +
  
  # Bar plot for steps
  geom_bar(aes(y = steps, fill = "Steps"), stat = "identity", alpha = 0.6, position = "dodge") +
  
  scale_y_continuous(name = "Heart Rate (bpm)",
                     sec.axis = sec_axis(~., name = "Steps")) + 
  labs(title = "Heart Rate and Steps (Binned in x-Minute Intervals)",
       x = "Time",
       y = "Heart Rate (bpm)",
       color = "Metrics") +
  facet_wrap(~ as.Date(bin_min), scales = "free_x") +
  theme_minimal() +
  scale_x_datetime(date_labels = "%H:%M", date_breaks = "1 hour") +  # Adjust time format and breaks
  scale_color_manual(values = c("Heart Rate" = "blue")) +
  scale_fill_manual(values = c("Steps" = "darkgreen"))

```

# heart v2 list
Returns a list of ECG records and Afib classification for a given period of time. To get the full ECG signal, use the Heart v2 - Get service.

```{r}

req <- httr::GET("https://wbsapi.withings.net/v2/heart?action=list", 
                 query = list(access_token = access_token,
                              startdate = UnixStartDate, 
                              enddate = UnixEndDate))

httr::stop_for_status(req)
out <- httr::content(req, as = "text", encoding = "utf-8")
out <- jsonlite::fromJSON(out)

out$status

out$body$offset

if (out$status == 0) {
        out$body$series <- jsonlite::flatten(out$body$series)
        out$body$series$timestamp <- as.POSIXct(out$body$series$timestamp, 
            tz = "", origin = "1970-01-01")
        out$body$series$modified <- as.POSIXct(out$body$series$modified, 
            tz = "", origin = "1970-01-01")
        out$body$series <- dplyr::arrange(out$body$series, startdate)
}

ECGList.DF<-out$body$series

# select the signal ids I want to examine from the ECGList.DF
ecg.signalIDs<-ECGList.DF$ecg.signalid

#write.csv(ECGList.DF,"c:/Documents and Settings/d.vitali/Downloads/ecg.signalIDs.csv")

```

## heart v2 get()
Provides the high frequency data of an ECG recording in micro-volt (μV).

needs ecg.signalID


```{r eval=FALSE}

# Fetch ECG data for all signal IDs
ecg_data_list <- fetch_ecg_signals(ecg.signalIDs, access_token)


library(ggplot2)

for (i in seq_along(ecg_data_list)) {
  # Get the current ECG dataframe
  ecg_df <- ecg_data_list[[i]]
  
  # Create and display the plot
  plot <- ggplot(ecg_df, aes(x = Timestamp, y = Value)) +
    geom_line() +
    labs(title = paste("ECG Signal for Signal ID", ecg.signalIDs[i]),
         x = "Timestamp (ms)",
         y = "muV") +
    theme_minimal()
  
  # Display the plot
  print(plot)
}



```

# sleep v2 get
```{r}

data_fields <-"hr,rr,snoring,sdnn_1,rmssd,mvt_score"
#data_fields<- ""
# > hr        : Heart Rate.
# > rr        : Respiration Rate.
# > snoring   : Total snoring time
# > sdnn_1    : Heart rate variability - Standard deviation of the NN over 1 minute
# > rmssd     : Heart rate variability - Root mean square of the successive 
#               differences over "a few seconds"
# > mvt_score : Track the intensity of movement in bed on a minute-by-minute basis. 
#               Only available for EU devices and devices under prescription.


startdate
enddate


# Fetch the sleep data for the week
All_sleep_data <- fetch_ALL_sleep_data(startdate, enddate, access_token, data_fields)

#write.csv(sleep.df,"c:/Documents and Settings/d.vitali/Downloads/sleep.csv")


```

## sleep v2 summary()
```{r}
data_fields <-"hr_average,hr_max,hr_min,lightsleepduration,wakeupcount"


req <- httr::GET("https://wbsapi.withings.net/v2/sleep?action=getsummary",
                 query = list(access_token = access_token,
                              startdateymd = startdate, 
                              enddateymd = enddate,
                              data_fields = data_fields))

out <- httr::content(req, as = "text", encoding = "utf-8")
out <- jsonlite::fromJSON(out)

out$status

if (out$status == 0) {
        out$body$series <- jsonlite::flatten(out$body$series)
        out$body$series$startdate <- as.POSIXct(out$body$series$startdate, 
            tz = "", origin = "1970-01-01")
        out$body$series$enddate <- as.POSIXct(out$body$series$enddate, 
            tz = "", origin = "1970-01-01")
        out$body$series <- dplyr::arrange(out$body$series, startdate)
    }

sleepSum.df<-out$body$series

#write.csv(sleepSum.df,"c:/Documents and Settings/d.vitali/Downloads/sleepsum.csv")


```
