---
title: "withingsAPI"
output: 
  github_document:
    toc: True
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r echo=F}
library("tidyr")
library("dplyr")
library("httr")
library("jsonlite")
```

```{r}
APIendpoint<-"https://wbsapi.withings.net"
AccessURL<-"https://account.withings.com/oauth2/token"
AuthURL <- "https://account.withings.com/oauth2_user/authorize2"
TokenURL<- "https://wbsapi.withings.net/v2/oauth2"
CallBackURL<-"http://localhost:1410/"
```

set working directory
```{r, echo=FALSE}
setwd("C:/Users/d.vitali/Desktop/Github/CRIISP-WP6/Sensors-API/Withings-API/")
```


# Load Necessary functions
```{r}
# this should read withings_func.R which contains a set of functions required in the pipeline
source("withings_func.R")

```

# Define Participant ID

```{r}
#set credentials for this participant
Participant_id<-"credentials/P_01.conf"
  
```


# Load Credentials
```{r}

source(Participant_id)
# this should read client_id, client_secret, and refresh_token as environment variables
```


# API setup

LOGIN as this participant: 

```{r}
# Main logic to check if refresh_token is empty and handle accordingly
if (refresh_token == "") {
  cat("refresh_token is empty. Requesting new tokens.\n")
  tokens <- request_new_tokens()
} else {
  cat("refresh_token found. Refreshing existing token.\n")
  tokens <- refresh_existing_token(refresh_token)
}

# Access tokens can now be used for further API calls
access_token <- tokens$access_token
refresh_token <- tokens$refresh_token

```


```{r}

req <- httr::GET("https://wbsapi.withings.net/v2/user?action=getdevice", 
                 query = list(access_token = access_token))

httr::stop_for_status(req)
out <- httr::content(req, as = "text", encoding = "utf-8")
out <- jsonlite::fromJSON(out)

out$status

devices.df<-out$body$devices

deviceid<-out$body$devices$deviceid[1] # just in case I need to filter by deviceid


```
# SET start date enddate

Set this to three arbitrary days where I know there is data
```{r}

startdate   <-"2024-07-08" 
enddate     <-"2024-07-08"
```

transform the time format of "startdate" in the format required by the different API calls
```{r}
## set time variable as unixtime
UnixStartDate <- as.POSIXct(paste(startdate, "00:00:00"), format = "%Y-%m-%d %H:%M:%S", tz = "UTC")
UnixStartDate<- as.numeric(unclass(UnixStartDate))

UnixEndDate <- as.POSIXct(paste(enddate, "23:59:59"), format = "%Y-%m-%d %H:%M:%S", tz = "UTC")
UnixEndDate <- as.numeric(unclass(UnixEndDate))

## set time variable as timezone
startdate_tz<-convert_date(startdate, "Date", tz)
enddate_tz<-convert_date(enddate, "Date", tz)

```

#API CALLS

## getmeas()
I think this one is pretty useless
```{r}
#11	Heart Pulse (bpm) - only for BPM and scale devices
#123	VO2 max is a numerical measurement of your body’s ability to consume oxygen (ml/min/kg).
#54	SP02 (%)
#135	QRS interval duration based on ECG signal
#136	PR interval duration based on ECG signal
#137	QT interval duration based on ECG signal
#138	Corrected QT interval duration based on ECG signal

meastypes<-"11"#,54,123,135,136,136,137,138"
req <- httr::GET("https://wbsapi.withings.net/measure?action=getmeas",
             query=list(access_token=access_token,
                        meastype = meastypes,
                        category = 1, # 1 for real measures 2 for user objectives
                        startdate = UnixStartDate,
                        enddate = UnixEndDate))


out <- httr::content(req, as = "text", encoding = "utf-8")

out <- jsonlite::fromJSON(out)

out$status

 if(out$status==0) {

    out$body$updatetime = as.POSIXct(out$body$updatetime, tz="", origin="1970-01-01")

    if( length(out$body$measuregrps) > 1) {
      out$body$measuregrps <- jsonlite::flatten(out$body$measuregrps)
      out$body$measuregrps$date <- as.POSIXct(out$body$measuregrps$date, tz="", origin="1970-01-01")

      measures <- dplyr::bind_rows(out$body$measuregrps$measures)

      out$body$measuregrps <- dplyr::select(out$body$measuregrps, -"measures")

      out$body$measuregrps <- dplyr::bind_cols(out$body$measuregrps, measures)
    }
  }
getmeasures.df<-out$body$measuregrps

#write.csv(getmeasures.df,"c:/Documents and Settings/d.vitali/Downloads/getmeasures.csv")


```

## getactivity()

```{r}
##

req <- httr::GET("https://wbsapi.withings.net/v2/measure?action=getactivity", 
                 query = list(access_token = access_token,
                              startdateymd = startdate_tz, 
                              enddateymd = enddate_tz,
                              offset = ""))
httr::stop_for_status(req)
out <- httr::content(req, as = "text", encoding = "utf-8")
out <- jsonlite::fromJSON(out)

out$status


activity.df<-out$body$activities

#write.csv(activity.df,"c:/Documents and Settings/d.vitali/Downloads/getactivity.csv")


```






## getintradayactivity()

```{r}

data_fields <-"steps,elevation,calories,distance,duration,heart_rate,spo2_auto"


req <- httr::GET("https://wbsapi.withings.net/v2/measure?action=getintradayactivity", 
                 query = list(access_token = access_token,
                              startdate = UnixStartDate, 
                              enddate   = UnixEndDate,
                              data_fields = data_fields))
out <- httr::content(req, as = "text", encoding = "utf-8")
out <- jsonlite::fromJSON(out)

out$status

if (out$status == 0) {
        out$body$series <- lapply(out$body$series, function(x) x[!sapply(x, 
            is.null)])
        out$body$series <- dplyr::bind_rows(out$body$series, 
            .id = "timestamp")
        out$body$series$timestamp <- as.POSIXct(as.numeric(out$body$series$timestamp), 
            tz = "", origin = "1970-01-01")
    }

intraDayActivity.df<-out$body$series

#write.csv(intraDayActivity.df,"c:/Documents and Settings/d.vitali/Downloads/getintraDayActivity.csv")

```


## getworkouts()
```{r}

#lastupdate  <-         #Timestamp for requesting data that were updated or 
                        #created after this date. Useful for data synchronization 
                        #between systems. Use this instead of startdateymd + enddateymd.

datafields<-"calories,intensity,manual_distance,manual_calories,hr_average,hr_min,hr_max,hr_zone_0,hr_zone_1,hr_zone_2,hr_zone_3,pause_duration,algo_pause_duration,spo2_average,steps,distance,elevation,pool_laps,strokes,pool_length"

req <- httr::GET("https://wbsapi.withings.net/v2/measure?action=getworkouts", 
                 query = list(access_token = access_token,
                              startdateymd = startdate, 
                              enddateymd   = enddate,
                              data_fields = datafields))
out <- httr::content(req, as = "text", encoding = "utf-8")
out <- jsonlite::fromJSON(out)

out$status

# Flatten the nested data
workouts.df <- out$body$series %>%
  unnest(cols = c(data)) %>%
  mutate(
    startdate = as.POSIXct(startdate, origin = "1970-01-01"),
    enddate = as.POSIXct(enddate, origin = "1970-01-01"),
    modified = as.POSIXct(modified, origin = "1970-01-01")
  )

#write.csv(workouts.df,"c:/Documents and Settings/d.vitali/Downloads/getworkouts.csv")


```

# heart v2 list
Returns a list of ECG records and Afib classification for a given period of time. To get the full ECG signal, use the Heart v2 - Get service.

If the ECG recordings have been taken with BPM Core, systole and diastole measurements will also be returned.
```{r}

req <- httr::GET("https://wbsapi.withings.net/v2/heart?action=list", 
                 query = list(access_token = access_token,
                              startdate = UnixStartDate, 
                              enddate = UnixEndDate))

httr::stop_for_status(req)
out <- httr::content(req, as = "text", encoding = "utf-8")
out <- jsonlite::fromJSON(out)

out$status

out$body$offset

if (out$status == 0) {
        out$body$series <- jsonlite::flatten(out$body$series)
        out$body$series$timestamp <- as.POSIXct(out$body$series$timestamp, 
            tz = "", origin = "1970-01-01")
        out$body$series$modified <- as.POSIXct(out$body$series$modified, 
            tz = "", origin = "1970-01-01")
        out$body$series <- dplyr::arrange(out$body$series, startdate)
}

ECGList.DF<-out$body$series

# select the signal id I want to examine from the ECGList.DF
ecg.signalID<-out$body$series$ecg.signalid[1]

#write.csv(ECGList.DF,"c:/Documents and Settings/d.vitali/Downloads/ecg.signalIDs.csv")

```

## heart v2 get()
Provides the high frequency data of an ECG recording in micro-volt (μV).

needs ecg.signalID


```{r eval=FALSE}

### signalid is the id of the ECG signal recording from heart() list

req <- httr::GET(url = "https://wbsapi.withings.net/v2/heart", 
        query = list(access_token = access_token, 
            action = "get", signalid = ecg.signalID))
httr::stop_for_status(req)
out <- httr::content(req, as = "text", encoding = "utf-8")
out <- jsonlite::fromJSON(out)

out$status

ECGraw.df<-process_data(out)

library(ggplot2)
ggplot(ECGraw.df, aes(x = Timestamp, y = Value)) +
  geom_line() +
  labs(title = "Line Plot of Values Over Time",
       x = "Timestamp",
       y = "muV") +
  theme_minimal()

#write.csv(ECGraw.df,"c:/Documents and Settings/d.vitali/Downloads/ecg.track.csv")

```


# sleep v2 get
```{r}

data_fields <-"hr,rr,snoring,sdnn_1,rmssd,mvt_score"
#data_fields<- ""
# > hr        : Heart Rate.
# > rr        : Respiration Rate.
# > snoring   : Total snoring time
# > sdnn_1    : Heart rate variability - Standard deviation of the NN over 1 minute
# > rmssd     : Heart rate variability - Root mean square of the successive 
#               differences over "a few seconds"
# > mvt_score : Track the intensity of movement in bed on a minute-by-minute basis. 
#               Only available for EU devices and devices under prescription.





req <- httr::GET("https://wbsapi.withings.net/v2/sleep?action=get", 
                 query = list(access_token = access_token,
                              startdate = UnixStartDate, 
                              enddate = UnixEndDate,
                              data_fields = data_fields))
httr::stop_for_status(req)
out <- httr::content(req, as = "text", encoding = "utf-8")
out <- jsonlite::fromJSON(out)

out$status

if (out$status == 0) {
    # Flatten the JSON data
    sleep.df <- jsonlite::flatten(out$body$series)
    
    # Convert startdate and enddate to POSIXct
    sleep.df$startdate <- as.POSIXct(sleep.df$startdate, tz = "", origin = "1970-01-01")
    sleep.df$enddate <- as.POSIXct(sleep.df$enddate, tz = "", origin = "1970-01-01")
    
    # Arrange by startdate
    sleep.df <- dplyr::arrange(sleep.df, startdate)
    
    # Separate heart rate columns
    sleep.hr_columns <- grep("^hr\\.", names(sleep.df), value = TRUE)
    
    # Gather heart rate data into long format
    sleep.df <- sleep.df %>%
        select(startdate, enddate, state, model, all_of(sleep.hr_columns)) %>%
        gather(key = "timestamp", value = "heart_rate", all_of(sleep.hr_columns)) %>%
        filter(!is.na(heart_rate))
    
    # Convert timestamp to POSIXct
    sleep.df$timestamp <- as.numeric(sub("hr\\.", "", sleep.df$timestamp))
    sleep.df$timestamp <- as.POSIXct(sleep.df$timestamp, tz = "", origin = "1970-01-01")
    
    # Create the main dataframe without heart rate columns
    main_df <- sleep.df %>%
        select(startdate, enddate, state, model)
}

sleep.df$state<-factor(sleep.df$state,
                       levels = c(0,1,2,3),
                       labels = c("Awake","LightSleep","DeepSleep","REM"))

#write.csv(sleep.df,"c:/Documents and Settings/d.vitali/Downloads/sleep.csv")


```

## sleep v2 summary()
```{r}
data_fields <-""


req <- httr::GET("https://wbsapi.withings.net/v2/sleep?action=getsummary",
                 query = list(access_token = access_token,
                              startdateymd = startdate, 
                              enddateymd = enddate))

out <- httr::content(req, as = "text", encoding = "utf-8")
out <- jsonlite::fromJSON(out)

out$status

if (out$status == 0) {
        out$body$series <- jsonlite::flatten(out$body$series)
        out$body$series$startdate <- as.POSIXct(out$body$series$startdate, 
            tz = "", origin = "1970-01-01")
        out$body$series$enddate <- as.POSIXct(out$body$series$enddate, 
            tz = "", origin = "1970-01-01")
        out$body$series <- dplyr::arrange(out$body$series, startdate)
    }

sleepSum.df<-out$body$series

#write.csv(sleepSum.df,"c:/Documents and Settings/d.vitali/Downloads/sleepsum.csv")


```
