---
title: "Fitbit API"
output: 
  github_document:
    toc: True
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
```{r echo=F}
library("fitbitr")
#library("reticulate")
#py_install("fitbit")
library("ggplot2")

```

## Use

### Callback URL

Note: when setting up an app, the Callback URL should be set to http://localhost:1410/. This is not supposed to be allowed, but it does seems to work. This is handy because allows for the callback to be (e.g.) on a laptop for testing.

```{r eval=FALSE}
#devtools::install_github("mrkaye97/fitbitr")

FITBIT_CLIENT_ID <- "23RBL4"
FITBIT_CLIENT_SECRET <- "5584bef963882c9134273b0f77574bb5"
CALLBACK_URL<-"http://localhost:1410/"
authorization_URI<-"https://www.fitbit.com/oauth2/authorize"
Token_Request_URI<-"https://api.fitbit.com/oauth2/token"

APP_NAME<-"sense1_test"

fitbitr_token <- generate_fitbitr_token(
  client_id = FITBIT_CLIENT_ID,
  client_secret = FITBIT_CLIENT_SECRET,
  callback = CALLBACK_URL,
  cache = TRUE
)




```
## Heart rate
```{r}

# Example date
date <- "2023-10-19"
start_date <- lubridate::today() - lubridate::weeks(1)
end_date <- lubridate::today()


heartR_6Hdata<-get_heart_rate_intraday(
  date = date, #lubridate::today(),
  detail_level = "1sec",#c("1sec", "1min", "5min", "15min"),
  start_time = "06:00:00",
  end_time = "12:00:00"
) 

heartR_1mindata<-get_heart_rate_intraday(
  date = date, #lubridate::today(),
  detail_level = "1sec",#c("1sec", "1min", "5min", "15min"),
  start_time = "17:00:00",
  end_time = "17:01:00"
) 


ggplot(heartR_6Hdata, aes(x=time, y=heart_rate)) +
  geom_line() + 
  xlab("")

 ggplot(heartR_1mindata, aes(x=time, y=heart_rate)) +
  scale_x_datetime(date_breaks = "2 sec", date_labels = "%OS")+
  #scale_x_datetime(minor_breaks = "2 sec")+
  geom_line() + 
  geom_point() +
  xlab(paste("1 minute from:","12:15:00"))



```

## STEPS and distance
```{r}

start_time <- "07:20:00"
end_time   <- "09:35:00"

distance<-get_distance_intraday(
date = date, #lubridate::today(),
detail_level = "1min", #c("1min", "5min", "15min"),
start_time = start_time,
  end_time = end_time
)
steps<-get_steps_intraday(
date = date,# lubridate::today(),
detail_level = "1min", #c("1min", "5min", "15min"),
start_time = start_time,
  end_time = end_time
)

elevation<-get_elevation_intraday(
date = date,# lubridate::today(),
detail_level = "1min", #c("1min", "5min", "15min"),
start_time = start_time,
  end_time = end_time
)


p_dist<-ggplot(distance, aes(x=time, y=distance*1000)) +
  geom_line() + 
  ylab("distance (m)")+
  xlab("time")

p_steps<-ggplot(steps, aes(x=time, y=steps)) +
  geom_line() + 
  xlab("time")

p__elev<-ggplot(elevation, aes(x=time, y=elevation, alpha = .4)) +
  geom_area() + 
  xlab("time")

p_dist+ geom_area(data = elevation, aes(x=time, y=elevation, alpha = .4))

p_dist + geom_line(data = steps, aes(x = time, y = steps), color = "red") +
  scale_y_continuous(
    name = "Distance (m)",
    sec.axis = sec_axis(~., name = "Steps")
  ) +
  scale_color_manual(values = c("blue", "red"))
  
```

```{r}

date<- "2023-10-19"
start_time <- "07:20:00"
end_time   <- "09:35:00"

timeserie<-get_intraday_time_series(
  resource = c("active-zone-minutes", "calories", "distance", "elevation", "floors",
    "heart", "steps"),
  date,
  detail_level,
  start_time,
  end_time
)

```


## HRV
Heart rate variability (HRV) is the physiological phenomenon of variation in the time interval between heartbeats. It is measured by the variation in the beat-to-beat interval.

The "Heart Rate Variability Details" files include 5 minutes granularity recordings of your HRV during a sleep. The description for the values of each row is as follows:

- timestamp: the start of the 5 minutes interval for which the following values were computed
- rmssd: "root mean square of successive differences" - the square root of the mean of the squares of the successive differences between adjacent beat-to-beat intervals
- coverage: the number of data points in the interval, multiplied by the mean beat-to-beat of the interval in seconds and divided by the number of seconds in the interval (300 seconds)
- low_frequency: measures long term variations in heart rate and reflects activity from both the sympathetic and parasympathetic branches
- high_frequency: measures short term variations in heart rate and captures parasympathetic activity

## Respiratory rate

The respiratory rate (or breathing rate) is the rate at which breathing occurs. This is usually measured in breaths per minute.

The "Respiratory Rate Summary" files include summaries of your respiratory rate during sleeps:

- timestamp: the wake time
- full_sleep_breathing_rate: the respiratory rate average for your entire sleep
- full_sleep_standard_deviation: measures the amount your respiratory rate variates during the entire sleep
- full_sleep_signal_to_noise: the signal-to-noise value for the entire sleep
- deep_sleep_breathing_rate: the respiratory rate average for your deep sleep periods
- deep_sleep_standard_deviation: measures the amount your respiratory rate variates during the deep sleep periods
- deep_sleep_signal_to_noise: the signal-to-noise value for the deep sleep periods
- light_sleep_breathing_rate: the respiratory rate average for your light sleep periods
- light_sleep_standard_deviation: measures the amount your respiratory rate variates during the light sleep periods
- light_sleep_signal_to_noise: the signal-to-noise value for the light sleep periods
- rem_sleep_breathing_rate: the respiratory rate average for your REM sleep periods
- rem_sleep_standard_deviation:measures the amount your respiratory rate variates during the REM sleep periods
- rem_sleep_signal_to_noise: the signal-to-noise value for the REM sleep periods